SNP Analysis

### 2018-07-16 SNP_Prediction_Analysis.Rmd
...

@@

There's a significant difference in the numbers of NDBE or HGD samples between the groups
```{r}
chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',])
chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',])

# Most progressors are HGD at timepoint 1
with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status)
```

@@

```{r, echo=F, warning=F}
qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
head(qcdata)

@@

```{r, echo=F, warning=F}

@@

@@

seg = ascat.output$segments_raw
seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA', subtitle='patient 403')

@@

seg = ascat.output$segments_raw
seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')

@@

```{r, warning=F, echo=F}

@@

```{r normals, fig.width=6, fig.height=8, eval=F}
nm = melt(normalSegs)
be = melt(mergedSegs[-nm.rows,])

@@

grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims)
)

@@

```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=T}
randNormal = sample(rownames(normalSegs), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

@@

randBE = '512_133R'
randBE = '1006_190V'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

@@

sa = subtract.arms(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd),
              adjust.cols(mergedArms[sub('\\.LogR','',randBE),,drop=F], z.arms.mean, z.arms.sd))[,1:589]
ggplot(melt(as.matrix(sa)), aes(Var1, value)) + geom_point()

@@

Using the mean bin values (merged during tiling).
```{r, warning=F}
# z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
# z.arms.mean[is.na(z.arms.mean)] = median(z.arms.mean,na.rm=T)
# 
# z.sd[is.na(z.sd)] = median(z.sd,na.rm=T)
# z.arms.sd[is.na(z.arms.sd)] = median(z.arms.sd,na.rm=T)

@@

grid.arrange(
  ggplot(melt(ms[sub('\\.LogR','',randBE),,drop=F]), aes(Var2, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
  ggplot(melt(subtract.arms(ms[sub('\\.LogR','',randBE),,drop=F],ma[sub('\\.LogR','',randBE),,drop=F])), aes(Var2, value)) +
    geom_point() + theme(axis.text.x=element_blank()) + labs(title='Arms subtracted col adjusted LRR, binned'),  
  top=randBE)
# So what if I don't subtract the arms?

@@

Normal variance: `r var(nm$value, na.rm=T)`

BE variance: `r var(be$value, na.rm=T)`

```{r echo=F}
cxN = score.cx(ns, 1)
arrayDfNorm = subtract.arms(ns, na)
#arrayDfNorm = cbind(ns, na)
arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))

@@

# grid.arrange(
#     ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + geom_point(data=melt(arrayDf[,589:634]+abs(mean(arrayDf[,589:634])*2)), col='red') + labs(title='SNP'),
#     ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
#       geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
#   nrow=1)

@@

## TEMP: I accidentally included normals by failing to account for difference in name
#preds = preds[-grep('_\\d+$', rownames(preds)),]

#qqnorm(allDf, main='Normal Q-Q plot, sWGS')
#qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')

>>>>

`r nrow(patient.info)` total endoscopies, `r pander(table(patient.info$Status), caption="Split between P and NP")`

Histology breakdown:

@@

There's a significant difference in the numbers of NDBE or HGD samples between the groups
```{r, echo=F}
pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')

# Most progressors are HGD at timepoint 1
pander( round(with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status), 3)*100, caption='% at T1')
```

@@

```{r, echo=F, warning=F, message=F}
qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
head(qcdata)

@@

```{r, echo=F, warning=F, fig.height=10, fig.width=8}

@@

summary.file = '~/Data/Annalise/NonP/summary_logR.txt'
smy = read.table(summary.file, sep = '\t', header = T)
smy = subset(smy, Copy.number <= 12)

a = as.data.frame(smy[,c('Copy.number','var.LRR')])
colnames(a) = c('x','y')
b = as.data.frame(smy[,c('Copy.number','median.LRR')])
colnames(b) = c('x','y')
b[b$x == 2,'y'] = 0

fitSD = lm(x~y, data=a)
autoplot(fitSD)
fitM = lm(x~y, data=b)
autoplot(fitM)

# Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
adjust.segraw<-function(segraw) {
  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
  segraw = subset(segraw, chr %in% c(1:22))

  segraw$adjLRR = NA
  for (cn in unique(segraw$total)) {
    rows = which(segraw$total == cn)
    values = seg[rows, 'medLRR', drop=T]

    #newM = mean(values)
    #newSD = sd(values)

    #if (cn != 2) {
      #newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
      #newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))

    #newV = mean(values)

    #}
    #segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))

    segraw[rows,'adjLRR'] = values * predict(fitM, newdata=cbind.data.frame('y'=values)) - cn)
  }  
segraw  
}

@@

seg = ascat.output$segments_raw
seg = adjust.segraw(seg)

m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)

seg$chr = factor(seg$chr, levels=c(1:22))
grid.arrange(
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, medLRR', subtitle=paste('patient',pt)),
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, adjLRR', subtitle=paste('patient',pt))
)

@@

seg = ascat.output$segments_raw
seg = adjust.segraw(seg)

m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)

seg$chr = factor(seg$chr, levels=c(1:22))
grid.arrange(
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, medLRR', subtitle=paste('patient',pt)),
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, adjLRR', subtitle=paste('patient',pt))
)

@@

```{r, warning=F, echo=F, message=F, fig.height=10, fig.width=8}

@@

```{r normals, message=F, fig.height=10, fig.width=8}
nm = melt(normalSegs)
be = melt(mergedSegs[-nm.rows,])

@@

grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric',subtitle=paste('variance=',var(nm$value),sep=''),x='') + 
    theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value),sep=''),x='') + 
    theme(axis.text.x = element_blank()) + ylim(lims)
)

@@

```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8}
randNormal = sample(rownames(normalSegs), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

@@

#randBE = '512_133R'
randBE = '1006_190V'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

@@

@@

Using the mean bin values (merged during tiling).
```{r, warning=F, message=F, fig.width=10, fig.height=8}
# z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
# z.arms.mean[is.na(z.arms.mean)] = median(z.arms.mean,na.rm=T)
# 
# z.sd[is.na(z.sd)] = median(z.sd,na.rm=T)
# z.arms.sd[is.na(z.arms.sd)] = median(z.arms.sd,na.rm=T)

@@

grid.arrange(
  ggplot(melt(ms[sub('\\.LogR','',randBE),,drop=F]), aes(Var2, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
  ggplot(melt(subtract.arms(ms[sub('\\.LogR','',randBE),,drop=F],ma[sub('\\.LogR','',randBE),,drop=F])), aes(Var2, value)) +
    geom_point() + theme(axis.text.x=element_blank()) + labs(title='Arms subtracted col adjusted LRR, binned'),  
  top=randBE)

@@

```{r echo=F,message=F, fig.width=10, fig.height=8}
cxN = score.cx(ns, 1)
arrayDfNorm = subtract.arms(ns, na)
#arrayDfNorm = cbind(ns, na)
arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))

@@

@@

qqnorm(allDf, main='Normal Q-Q plot, sWGS')
qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')
